The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-05-14 Paper/jaccardbaetica.tex
last review draft

Dressel 20 was the amphora most stamped during the Roman Empire \citep[18]{millet_anforas_1998}. The  period of use is associated with a large amphora production for almost three centuries.

Most stamps display a large activity of production that it can be difficult for specifying an accurate chronology in some cases. They have been frequently dated with the consular dating by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
However, the chronology of stamps by using \textit{tituli picti} could be also biased in other sites taking into account that amphorae were deposited after being produced and consumed.  
\maria{no sé cómo explicarlo mejor ni como citarlo porque no hay citas}

@@

The goal of this study is analysing the effect of the production patterns between different centres. We are specially interested in identifying links between producers and consumption centres using amphoric stamps. To do that, we use the CEIPAC database to collect the stamps from different places. The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}. Therefore, this study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps. A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps. Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.

>>>>

Dressel 20 was the amphora most stamped during the Roman Empire \citep[18]{millet_anforas_1998}. The usage period is associated with a large amphora production for almost three centuries.

Most stamps display a large activity of production that it can be difficult for specifying an accurate chronology in some cases. They have been frequently dated with the consular dating by studying \textit{tituli picti} found in \textit{Monte Testaccio} \citep{Testaccio1, berni_millet_epigrafianforica_2008}.
However, the chronology of stamps could be also biased in other sites taking into account that amphorae were deposited after being produced and consumed.  
\maria{no sé cómo explicarlo mejor ni como citarlo porque no hay citas}

@@

The goal of this study is exploring the effect of the production patterns between different centres. We are specially interested in identifying links between producers and consumption centres by using amphoric stamps. To do that, we use the CEIPAC database to collect the stamps from different places. The CEIPAC dataset contains over 50.000 of epigraphy records found in amphorae, mostly from \textit{Monte Testaccio}. Therefore, this study proposes a robust baseline to explore the distribution of Baetican olive oil production by computing the spatial correlation between stamps. A way to analyse is to use a quantitative framework to measure the similarity between amphoric stamps. Here we use an ecological approach based on three steps: a) to detect similarities between stamp codes, b) to explore a potentially spatial correlation and c) to establish a correlation between similarity of stamps and spatial distance.
